---
title: "wk7"
author: "Tim"
date: "2021/11/25"
output: html_document
---

```{r}
#library a bunch of packages we may (or may not) use - install them first if not installed already. 
install.packages('mapview')
install.packages('car')

library(tidyverse)
library(tmap)
library(geojsonio)
library(plotly)
library(rgdal)
library(broom)
library(mapview)
library(crosstalk)
library(sf)
library(sp)
library(spdep)
library(car)
library(fs)
library(janitor)
```

```{r}
#download a zip file containing some boundaries we want to use

download.file("https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip", 
              destfile="data/statistical-gis-boundaries-london.zip")

```
```{r}
listfiles<-dir_info(here::here("data")) %>%
  dplyr::filter(str_detect(path, ".zip")) %>%
  dplyr::select(path)%>%
  pull()%>%
  #print out the .gz file
  print()%>%
  as.character()%>%
  utils::unzip(exdir=here::here("data"))
```

```{r}
#look what is inside the zip

Londonwards<-dir_info(here::here("data", 
                                 "statistical-gis-boundaries-london", 
                                 "ESRI"))%>%
  #$ means exact match
  dplyr::filter(str_detect(path, 
                           "London_Ward_CityMerged.shp$"))%>%
  dplyr::select(path)%>%
  pull()%>%
  #read in the file in
  st_read()
```

```{r}
#read in some attribute data
LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               col_names = TRUE, 
                               locale = locale(encoding = 'Latin1'))
```

```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>% 
  clean_names() %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```

Deal with NA values which change data type from numeric to character

```{r}
LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               na = c("", "NA", "n/a"), 
                               locale = locale(encoding = 'Latin1'), 
                               col_names = TRUE)
```

```{r}
Datatypelist <- LondonWardProfiles %>% 
  clean_names() %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```


```{r}
#merge boundaries and data
LonWardProfiles <- Londonwards%>%
  left_join(.,
            LondonWardProfiles, 
            by = c("GSS_CODE" = "New code"))

#let's map our dependent variable to see if the join has worked:
tmap_mode("plot")
qtm(LonWardProfiles, 
    fill = "Average GCSE capped point scores - 2014", 
    borders = NULL,  
    fill.palette = "Blues")
```

more data, school data

```{r}
#might be a good idea to see where the secondary schools are in London too
london_schools <- read_csv("https://data.london.gov.uk/download/london-schools-atlas/57046151-39a0-45d9-8dc0-27ea7fd02de8/all_schools_xy_2016.csv")

#from the coordinate values stored in the x and y columns, which look like they are latitude and longitude values, create a new points dataset
lon_schools_sf <- st_as_sf(london_schools, 
                           coords = c("x","y"), 
                           crs = 4326)

lond_sec_schools_sf <- lon_schools_sf %>%
  dplyr::filter(PHASE=="Secondary")

tmap_mode("plot")
qtm(lond_sec_schools_sf)
```

```{r}
q <- qplot(x = `Unauthorised Absence in All Schools (%) - 2013`, 
           y = `Average GCSE capped point scores - 2014`, 
           data=LonWardProfiles)

#plot with a regression line - note, I've added some jitter here as the x-scale is rounded
#ggplot -> stat_smooth -> lm(means linear model)

q + stat_smooth(method="lm", se=FALSE, size=1) + 
  geom_jitter()
```


Below is the code that could be used to draw the blue line in our scatter plot. Note, the tilde ~ symbol means “is modelled by”.
```{r}
#run the linear regression model and store its outputs in an object called model1
Regressiondata<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014, 
                unauthorised_absence_in_all_schools_percent_2013)

#now model
model1 <- Regressiondata %>%
  lm(average_gcse_capped_point_scores_2014 ~
               unauthorised_absence_in_all_schools_percent_2013,
     data=.)
```

```{r}
#show the summary of those outputs
summary(model1)
```

```{r}
library(broom)
tidy(model1)
```

```{r}
#also from broom, glance can get a bit summary infor
glance(model1)

```

```{r}
install.packages('tidypredict')
library(tidypredict)
Regressiondata %>%
  tidypredict_to_column(model1)
```

7.5.6 Assumption 1 - There is a linear relationship between the dependent and independent variables
```{r}
# use Janitor to clean up the names.

LonWardProfiles <- LonWardProfiles %>%
  clean_names()

#let's check the distribution of these variables first

ggplot(LonWardProfiles, aes(x=average_gcse_capped_point_scores_2014)) + 
  geom_histogram(aes(y = ..density..),
                 binwidth = 5) + 
  geom_density(colour="blue", 
               size=1, 
               adjust=1)
```
```{r}
ggplot(LonWardProfiles, aes(x=unauthorised_absence_in_all_schools_percent_2013)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 0.1) + 
  geom_density(colour="red",
               size=1, 
               adjust=1)
```

Contrast with the median house price
```{r}
ggplot(LonWardProfiles, aes(x=median_house_price_2014)) + 
  geom_histogram()
```

Example for bad linear relationship

```{r}
qplot(x = median_house_price_2014, 
      y = average_gcse_capped_point_scores_2014, 
      data=LonWardProfiles)
```

log transfoirmation
```{r}
ggplot(LonWardProfiles, aes(x=log(median_house_price_2014))) + 
  geom_histogram()
```

Fortunately in R, we can use the symbox() function in the car package to try a range of transfomations along Tukey’s ladder:
```{r}
car::symbox(~median_house_price_2014, 
       LonWardProfiles, 
       na.rm=T,
       powers=seq(-3,3,by=.5))
```

Observing the plot above, it appears that raising our house price variable to the power of -1 should lead to a more normal distribution:

```{r}
ggplot(LonWardProfiles, aes(x=(median_house_price_2014)^-1)) + 
  geom_histogram()
```

```{r}
qplot(x = (median_house_price_2014)^-1, 
      y = average_gcse_capped_point_scores_2014,
      data=LonWardProfiles)
```

comparing with log tran

```{r}
qplot(x = log(median_house_price_2014), 
      y = average_gcse_capped_point_scores_2014, 
      data=LonWardProfiles)
```

7.5.7 Assumption 2 - The residuals in your model should be normally distributed


use broom::augment to access the residual data from model1
```{r}
#save the residuals into your dataframe

model_data <- model1 %>%
  augment(., Regressiondata)

#plot residuals
model_data%>%
dplyr::select(.resid)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

7.5.8 Assumption 3 - No Multicolinearity in the independent variables

```{r}
Regressiondata2<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014)

model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), data = Regressiondata2)

#show the summary of those outputs
tidy(model2)
```
```{r}
glance(model2)

```

```{r}
#and for future use, write the residuals out
model_data2 <- model2 %>%
  augment(., Regressiondata2)

# also add them to the shapelayer
LonWardProfiles <- LonWardProfiles %>%
  mutate(model2resids = residuals(model2))
```

multicolinearity
```{r}
install.packages('corrr')
library(corrr)

Correlation <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014) %>%
  mutate(median_house_price_2014 =log(median_house_price_2014))%>%
    correlate() %>%
  # just focus on GCSE and house prices
  focus(-average_gcse_capped_point_scores_2014, mirror = TRUE) 


#visualise the correlation matrix
rplot(Correlation)
```
Variance Inflation Factor (VIF)
Another way that we can check for Multicolinearity
```{r}
vif(model2)
```

```{r}
position <- c(10:74)

Correlation_all<- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(position)%>%
    correlate()

rplot(Correlation_all)
```

```{r}
position <- c(10:74)

Correlation_all<- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(position)%>%
    correlate()

rplot(Correlation_all)
```

7.5.9 Assumption 4 - Homoscedasticity 侯某西打死踢死踢
Homoscedasticity means that the errors/residuals in the model exhibit constant / homogenous variance, if they don’t, then we would say that there is hetroscedasticity present. 

```{r}
#print some model diagnositcs. 
par(mfrow=c(2,2))    #plot to 2 by 2 array
plot(model2)
```

7.5.10 Assumption 5 - Independence of Errors
This assumption simply states that the residual values (errors) in your model must not be correlated in any way.
If they are, then they exhibit autocorrelation which suggests that something might be going on in the background that we have not sufficiently accounted for in our model.

a regression model on data that do not have explicit space or time dimensions - the standard test for autocorrelation would be the Durbin-Watson test.
A value greater than 2 suggesting negative autocorrelation and and value of less than 2 indicating postitve autocorrelation.
```{r}
#run durbin-watson test
DW <- durbinWatsonTest(model2)
tidy(DW)
```

spatial-autocorrelation.

map first
```{r}
#now plot the residuals
tmap_mode("view")
#qtm(LonWardProfiles, fill = "model1_resids")

tm_shape(LonWardProfiles) +
  tm_polygons("model2resids",
              palette = "RdYlBu") +
tm_shape(lond_sec_schools_sf) + tm_dots(col = "TYPE")
```
above map: blue close to blue, question very big. need to calculate spatial autocorr to decide

```{r}
#calculate the centroids of all Wards in London
coordsW <- LonWardProfiles%>%
  st_centroid()%>%
  st_geometry()

plot(coordsW)
```

```{r}
#Now we need to generate a spatial weights matrix 
#(remember from the lecture a couple of weeks ago). 
#We'll start with a simple binary matrix of queen's case neighbours

LWard_nb <- LonWardProfiles %>%
  poly2nb(., queen=T)

#or nearest neighbours
knn_wards <-coordsW %>%
  knearneigh(., k=4)

LWard_knn <- knn_wards %>%
  knn2nb()

#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
```

```{r}
plot(LWard_knn, st_geometry(coordsW), col="blue")

```


```{r}
#create a spatial weights matrix object from these weights

Lward.queens_weight <- LWard_nb %>%
  nb2listw(., style="W")

Lward.knn_4_weight <- LWard_knn %>%
  nb2listw(., style="W")
```

```{r}
Queen <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.queens_weight)%>%
  tidy()
Queen
```

```{r}
Nearest_neighbour <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()
Nearest_neighbour
```

Both have 0.2something, meaning there is a weak spatial autocorrelation 


7.6 Spatial Regression Models
7.6.1.1 The Spatial Lag (lagged dependent variable) model

```{r}
#Original Model
model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), data = LonWardProfiles)

tidy(model2)
```


```{r}
install.packages('spatialreg')
library(spatialreg)

slag_dv_model2_queen <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_nb, style="C"), 
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_queen)
```
```{r}
glance(slag_dv_model2_queen)

```
Rho is our spatial lag that measures the variable in the surrounding spatial areas as defined by the spatial weight matrix. We use this as an extra explanatory variable to account for clustering (identified by Moran’s I)


```{r}
t<-summary(slag_dv_model2_queen)

```

```{r}
#run a spatially-lagged regression model
slag_dv_model2_knn4 <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_knn, 
                        style="C"), 
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_knn4)
```

```{r}
#write out the residuals

LonWardProfiles <- LonWardProfiles %>%
  mutate(slag_dv_model2_knn_resids = residuals(slag_dv_model2_knn4))

KNN4Moran <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(slag_dv_model2_knn_resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

KNN4Moran
```

7.6.1.2 The Spatial Error Model

If there is no spatial autocorrelation in the residuals, we would expect the  
λ parameter to = 0

```{r}
sem_model1 <- errorsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles,
               nb2listw(LWard_knn, style="C"), 
               method = "eigen")

tidy(sem_model1)
```

7.6.2 Key advice

The Lagrange multiple tests are within the function lm.LMtests from the package spdep:
LMerr is the spatial error model test
LMlag is the lagged test

```{r}
library(spdep)

Lward.queens_weight_ROW <- LWard_nb %>%
  nb2listw(., style="W")

lm.LMtests(model2, Lward.queens_weight_ROW, test = c("LMerr","LMlag","RLMerr","RLMlag","SARMA"))
```


7.6.3 Which model to use
Usually you will run OLS regression first then look for spatial autocorrelation of the residuals (Moran’s I).

Once at this stage you need to make a decision about the model:

Is it a global model (error / lag) or a local model (GWR)?
Can a single model (error/lag) be fitted to the study area?
Is the spatial autocorrelation a problem (error) or showing local trends (GWR)?




GWR runs a local regression model for adjoining spatial units and shows how the coefficients can vary over the study area.


```{r}
extradata <- read_csv("https://www.dropbox.com/s/qay9q1jwpffxcqj/LondonAdditionalDataFixed.csv?raw=1")

#add the extra data too
LonWardProfiles <- LonWardProfiles%>%
  left_join(., 
            extradata, 
            by = c("gss_code" = "Wardcode"))%>%
  clean_names()

#print some of the column names
LonWardProfiles%>%
  names()%>%
  tail(., n=10)
```


7.6.5 Extending your regression model - Dummy Variables

```{r}
p <- ggplot(LonWardProfiles, 
            aes(x=unauth_absence_schools11, 
                y=average_gcse_capped_point_scores_2014))
p + geom_point(aes(colour = inner_outer)) 
```

```{r}
#first, let's make sure R is reading our InnerOuter variable as a factor
#see what it is at the moment...
isitfactor <- LonWardProfiles %>%
  dplyr::select(inner_outer)%>%
  summarise_all(class)

isitfactor
```


```{r}
# change to factor

LonWardProfiles<- LonWardProfiles %>%
  mutate(inner_outer=as.factor(inner_outer))

#now run the model
model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014) + 
               inner_outer, 
             data = LonWardProfiles)
 
tidy(model3)
```
 In this case, it is telling us that living in a Ward in outer London will improve your average GCSE score by 10.93 points, on average, compared to if you lived in Inner London. The R-squared has increased slightly, but not by much.

```{r}
contrasts(LonWardProfiles$inner_outer)

```

```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(inner_outer = relevel(inner_outer, 
                               ref="Outer"))

model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014) + 
               inner_outer, 
             data = LonWardProfiles)

tidy(model3)
```
Above: now inner is the coefficient now.



7.7 Task 3 - Spatial Non-stationarity and Geographically Weighted Regression Models (GWR)```{r}


```{r}
myvars <- LonWardProfiles %>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014,
         rate_of_job_seekers_allowance_jsa_claimants_2015,
         percent_with_level_4_qualifications_and_above_2011,
         inner_outer)

#check their correlations are OK
Correlation_myvars <- myvars %>%
  st_drop_geometry()%>%
  dplyr::select(-inner_outer)%>%
  correlate()

#run a final OLS model
model_final <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = myvars)

tidy(model_final)
```

```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(model_final_res = residuals(model_final))

par(mfrow=c(2,2))
plot(model_final)
```

```{r}
qtm(LonWardProfiles, fill = "model_final_res")

```

```{r}
final_model_Moran <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model_final_res)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

final_model_Moran
```

How scholars from Bristol do GWR
```{r}
install.packages('spgwr')
library(spgwr)

coordsW2 <- st_coordinates(coordsW)

LonWardProfiles2 <- cbind(LonWardProfiles,coordsW2)

GWRbandwidth <- gwr.sel(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = LonWardProfiles2, 
                        coords=cbind(LonWardProfiles2$X, LonWardProfiles2$Y),
                  adapt=T)
```

```{r}
GWRbandwidth
```
The optimal bandwidth is about 0.015 meaning 1.5% of all the total spatial units should be used for the local regression based on k-nearest neighbours. Which is about 9 of the 626 wards.
 To change this we would set the argument gweight = gwr.Gauss in the gwr.sel() function — gwr.bisquare() is the other option.

To conclude, we can:

  set the bandwidth in gwr.sel() automatically using:
  the number of neighbors
  a distance threshold
Or, we can set it manually in gwr() using:
  a number of neighbors
  a distance threshold



```{r}

gwr.model = gwr(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = LonWardProfiles2, 
                coords=cbind(LonWardProfiles2$X, LonWardProfiles2$Y), 
                adapt=GWRbandwidth,
                #matrix output
                hatmatrix=TRUE,
                #standard error
                se.fit=TRUE)

#print the results of the model
gwr.model
```

Coefficient ranges can also be seen for the other variables and they suggest some interesting spatial patterning. To explore this we can plot the GWR coefficients for different variables. 

```{r}
results <- as.data.frame(gwr.model$SDF)
names(results)
```

```{r}
#attach coefficients to original SF
LonWardProfiles2 <- LonWardProfiles %>%
  mutate(coefUnauthAbs = results$unauthorised_absence_in_all_schools_percent_2013,
         coefHousePrice = results$log.median_house_price_2014.,
         coefJSA = rate_of_job_seekers_allowance_jsa_claimants_2015,
         coefLev4Qual = percent_with_level_4_qualifications_and_above_2011)
```

```{r}
tm_shape(LonWardProfiles2) +
  tm_polygons(col = "coefUnauthAbs", 
              palette = "RdBu", 
              alpha = 0.5)
```
```{r}
#run the significance test
sigTest = abs(results$"log.median_house_price_2014.")-2 * results$"log.median_house_price_2014._se"


#store significance results
LonWardProfiles2 <- LonWardProfiles2 %>%
  mutate(GWRUnauthSig = sigTest)
```
If this is greater than zero (i.e. the estimate is more than two standard errors away from zero), it is very unlikely that the true value is zero, i.e. it is statistically significant (at nearly the 95% confidence level)


```{r}
tm_shape(LonWardProfiles2) +
  tm_polygons(col = "GWRUnauthSig", 
              palette = "RdYlBu")
```

